---
title: "Captstone Analysis"
output: html_document
date: "2023-12-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Case Study 1: 
Notes: setting up my R environment by loading the 'tidyverse','dplyr', 'ggplot2' and 'lubridate' packages
```{r Uploading packages for Analysis}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(lubridate)
```
##I began by importing data from 11-2022 to 10-2023 to get a years worth of data to analyze

```{r Import data}
tripdata_2022_11 <- read.csv("C:/Users/ejspe/OneDrive/Desktop/Capstone_Cycles/202211-divvy-tripdata.csv")
tripdata_2022_12 <- read.csv("C:/Users/ejspe/OneDrive/Desktop/Capstone_Cycles/202212-divvy-tripdata.csv")
tripdata_2023_01 <- read.csv("C:/Users/ejspe/OneDrive/Desktop/Capstone_Cycles/202301-divvy-tripdata.csv")
tripdata_2023_02 <- read.csv("C:/Users/ejspe/OneDrive/Desktop/Capstone_Cycles/202302-divvy-tripdata.csv")
tripdata_2023_03 <- read.csv("C:/Users/ejspe/OneDrive/Desktop/Capstone_Cycles/202303-divvy-tripdata.csv")
tripdata_2023_04 <- read.csv("C:/Users/ejspe/OneDrive/Desktop/Capstone_Cycles/202304-divvy-tripdata.csv")
tripdata_2023_05 <- read.csv("C:/Users/ejspe/OneDrive/Desktop/Capstone_Cycles/202305-divvy-tripdata.csv")
tripdata_2023_06 <- read.csv("C:/Users/ejspe/OneDrive/Desktop/Capstone_Cycles/202306-divvy-tripdata.csv")
tripdata_2023_07 <- read.csv("C:/Users/ejspe/OneDrive/Desktop/Capstone_Cycles/202307-divvy-tripdata.csv")
tripdata_2023_08 <- read.csv("C:/Users/ejspe/OneDrive/Desktop/Capstone_Cycles/202308-divvy-tripdata.csv")
tripdata_2023_09 <- read.csv("C:/Users/ejspe/OneDrive/Desktop/Capstone_Cycles/202309-divvy-tripdata.csv")
tripdata_2023_10 <- read.csv("C:/Users/ejspe/OneDrive/Desktop/Capstone_Cycles/202310-divvy-tripdata.csv")


```

##Combine all the data sets so all the data can be analyzed together

```{r Combine Excel Sheets}


trips_whole_yr <-bind_rows(tripdata_2022_11, tripdata_2022_12,tripdata_2023_01,
                           tripdata_2023_02, tripdata_2023_03,tripdata_2023_04, tripdata_2023_05, tripdata_2023_06,
                           tripdata_2023_07,tripdata_2023_08,tripdata_2023_09, tripdata_2023_10)
```
##check if combined files show same column names. This is so it is easier to analyze
```{r Check to see if data is synced}
colnames(trips_whole_yr)# 

```
##Created new columns to show the day of week and months
```{r Create New }
trips_whole_yr_rev$month <- format(as.Date(trips_whole_yr_rev$date), "%m")
trips_whole_yr_rev$day_of_week <- format(as.Date(trips_whole_yr_rev$day_of_week), "%A")
```


#Analyze 
##Assigning the day to each day in week
```{r}
trips_whole_yr_rev$day_of_week <- 
  ordered(trips_whole_yr_rev$day_of_week, levels = c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday'))

trips_whole_yr_rev %>%
  group_by(member_casual, day_of_week) %>%
  summarise(number_of_ride = n()) %>%
  arrange(day_of_week)
```



```{r}
trips_whole_yr_rev$month <- 
  ordered(trips_whole_yr_rev$month, levels = c('11', '12', '01', '02', '03', '04', '05','06','07','08','09','10'))

trips_whole_yr_rev %>%
  group_by(member_casual, month) %>%
  summarise(number_of_ride = n(),) %>%
  arrange(month) %>%print(n=27)
```


## seeing the difference between how many casual and members ride at different months of the year

```{r}
trips_whole_yr_rev %>%
  group_by(member_casual, month) %>%
  summarise(number_of_ride = n()) %>%
  ggplot(mapping=aes(x= month,y=number_of_ride,fill= member_casual,label = number_of_ride)) +geom_bar(position = "dodge", stat = "identity")+  geom_text(size = 3, position = position_stack(vjust = 0.25))

```

## seeing the difference between how many casual and members ride at different times of the week

```{r}
trips_whole_yr_rev %>%
  group_by(member_casual, day_of_week) %>%
  summarise(number_of_ride = n()) %>%
  ggplot(mapping=aes(x= day_of_week,y=number_of_ride,fill= member_casual,label = number_of_ride)) +geom_bar(position = "dodge", stat = "identity") +geom_text(size = 3, position = position_stack(vjust = 0.25))
```

 

## seeing how long members ride in comparison to casual in week

```{r}
trips_whole_yr_rev %>%
  group_by(member_casual, day_of_week) %>%
  summarise(average_ride_length = n()) %>%
  ggplot(mapping=aes(x= day_of_week,y=average_ride_length,fill= member_casual,label = average_ride_length)) 
+geom_bar(position = "dodge", stat = "identity") +geom_text(size = 3, position = position_stack(vjust = 0.25))
```


## seeing how long members ride in comparison to casual each month
```{r}
trips_whole_yr_rev %>%
  group_by(member_casual, month) %>%
  summarise(average_ride_length = n()) %>%
  ggplot(mapping=aes(x= month,y=average_ride_length,fill= member_casual, label = average_ride_length)) 
+geom_bar(position = "dodge", stat = "identity") +geom_text(size = 3, position = position_stack(vjust = 0.25))
```









## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
